A Motion Blur Resilient Fiducial For Quadcopter Imaging

### 2014-06-13 WACV/paper.tex
Proof reading changes

In order to be effective, fiducials (or simply, fiducials) need
to be easily detected in the scene. A variety of fiducials have
been proposed~\cite{NaimarkF02,ARToolkit02,Fiala05,Pitag13,runetag11}
in the literature.  These take the form of binary codes arranged into
rectangular grids~\cite{ARToolkit02,Fiala05} or other geometric
primitives arranged in predefined spatial
patterns~\cite{NaimarkF02,Pitag13,runetag11}.  Figure \ref{fig:teaser}
shows the popular ARTag \cite{Fiala05} fiducial as possibly seen from
a quadcopter. A problem for existing fiducials is that low-cost
quadcopters often exhibit very quick and erratic physical movements
that result in motion blur in the quadcopter's onboard camera. This
motion blur has an adverse effect on the recognition of fiducial
markers. This can be seen in Figure \ref{fig:teaser}--(b) where the
ARTag fiducial cannot be recognized due to motion blur. This is not
too surprising as most existing fiducials are not designed to handle
motion blur.

@@

{\textbf{Tracking}}~~Fiducial detection between successive video
frames can be considered as a tracking problem where the tracked
object is the fiducial.  There is a very large body of research
dedicated to tracking and interested readers are referred
to~\cite{Yilmaz:2006} for a good survey.

@@

Here we examine the the performance of two popular fiducials under
motion blur.  Specifically we examine ARTags ~\cite{Fiala05} and
PiTags~\cite{Pitag13} given their differences in geometric design and
the availability of an API to develop applications to recognize the tags.
\footnote{RUNE-tag~\cite{runetag11} currently does not
provide access to an implementation to generate or recognize its tags. 
The circular data matrix~\cite{NaimarkF02} is available as a
commercial product, however it requires proprietary hardware.}

@@

%\noindent Details of each of the steps are as follows:\\
\textbf{Gabor filter}~~A 2D Gabor filter is a Gaussian kernel function
modulated by a sinusoidal plane wave~\cite{Kruizinga:2002}. It is used to find
high gradient patches. In our case, it is used to detect portions
of the circular fiducials that were not affected by the blur.
We applied the Gabor filter in eight
different orientations ($\theta = 0, 45, 90, \ldots, 225, 270, 315$).  The
following parameters were used for creating each Gabor kernel: $\lambda$ (wavelength) $= 8$, $\gamma$
(aspect ratio) $= 0.5$, $\sigma$ (spread) $= 0.56\lambda$, $\psi$
(phase angle) $= 0$ (for real part), $\pi/2$ (for imaginary part).
Then $\ell^2$-norm of outputs along all orientations is calculated and finally
$\ell^2$-normalized image is thresholded with threshold set to $0.4$ (on scale of zero to
one).  For further details to the Gabor filter, see~\cite{Kruizinga:2002}.
%% MBS - is the Kruizinga filter reasonable to cite?

\textbf{Clustering}~~The binarized Gabor filter responses are
treated as a set of connected components in the image.   Clustering
is used to find components that are located in a close spatial region.  We do
this via hierarchical clustering~\cite{ALGLIB} using unweighted
average linkage with a distance threshold set to 150.

\textbf{PCA}~~For each clustered set of components, we apply
PCA to determine the dominant direction.  This is done by examining
the orientation of the first principal component.  We then extract an
intensity profile patch in the input image along this direction as it
extends through the bounding box of the cluster. The signature of this
profile will be used to identify the code.

@@

\textbf{Classification}~~As mentioned above, a small image patch
containing the intensity profile of the fiducial is used to identify
which of the many fiducials might be present and is represented by a
code.  We found that a training-based method using the k nearest
neighbor (k-NN) technique gave better results than trying to find
binary code directly by determining the presence (or absence) of ring
at particular positions in the image patch. This required
training-data which was easy to generate. A synthetic fiducial
is blurred along 36 orientations (0, 10, 20, \ldots , 350) with blur
scale set to 40, and the intensity profile along the first principal
component from every output is taken as training data for that
fiducial. Figure~\ref{fig:training_data} shows the process of
creating training data for fiducial with binary code ``01'' embedded
in it.  For a query image patch, we normalize the intensity range, and
then compare this information against the training data. The class
label from the closest top $K=5$ images in the training data is used
to label the patch.

@@

Our system has also been tested on images containing multiple fiducial
patterns in the same frame. Our algorithm can successfully detected
all fiducials as well as correctly classified them as shown in
Figure \ref{fig:output_all}.

@@

Qualitatively we see in Figure~\ref{fig:blur_maximum} that despite
even more blur than earlier described (50 units or more) detection is
still feasible and practical.  Quantitatively, the comparison of
recognition rate is shown in Figure~\ref{fig:recognition_rate}.
Even under more severe blur, of 65 pixels, the algorithm is
successful.  Our recognition is 100\% for all codes except the ``00''
which is undetectable after 50 pixels blur (which is still
significantly better than ARTag).  Reasons for the lower performance
when``00'' tags are present is discussed in
Section~\ref{sec:discussion}.

@@

In another setup we arranged our fiducials on four sides of a box and
revolved the quadcopter around the box. Later, we repeated the process
by replacing our fiducial by ARTags. Figure~\ref{fig:setup} some
frames from this sequence as well as the performance of both
fiducials. The overall detection rate of our tags is 90\% while that
of ARTags is only around 60\%.

@@

We also compare our approach with tracking designed for blurred input
scenes.  We have used four image sequences (consisting of around 1000
frames each). Each sequence contains different fiducials so that we
are able to contrast the performance of BLUT~\cite{Wu:2011} with the
proposed method.  From the images in the top rows of
Figure~\ref{fig:BLUT_compare_00} and Figure~\ref{fig:BLUT_compare_01},
we can see that BLUT is able to track the fiducial when the position
of fiducial does not change too much in successive frames. Also, it
can be seen that once BLUT loses track of the fiducial, it is not
able to recover. Since our approach detects the code in each frame,
large changes in the patterns position is not an issue.  Some of our
detection results are also shown in Figure~\ref{fig:BLUT_compare_00}
and Figure~\ref{fig:BLUT_compare_01}.

@@

We found that sometimes our detection algorithm fails to recognize the
``00'' fiducial, when there is severe blur.  The is because when there
is too much blur, the innermost ring's response in the Gabor output is
too low and not detected properly.  This problem can be resolved by
increasing the radius of innermost ring to reduce the effect of blur
on the innermost ring.
%\noindent\textbf{Pose Estimation}~~

@@

In terms of numbers of fiducials, we are able to generate less markers
than ARTag. Many applications in robotics (e.g., quadcopter navigation)
may not require simultaneous use of a large number of fiducials.  Most
of the time, it is sufficient to have 4-6 different
fiducials. Nevertheless, we may be able to generate a larger number of
fiducials by using color backgrounds. This is an area of interest for
further investigation.

>>>>

In order to be effective, fiducial markers (or simply, fiducials) need
to be easily detected in the scene. A variety of fiducials have
been proposed~\cite{NaimarkF02,ARToolkit02,Fiala05,Pitag13,runetag11}
in the literature.  These take the form of binary codes arranged into
rectangular grids~\cite{ARToolkit02,Fiala05} or other geometric
primitives arranged in predefined spatial
patterns~\cite{NaimarkF02,Pitag13,runetag11}.  Figure \ref{fig:teaser}
shows the popular ARTag \cite{Fiala05} fiducial as possibly seen from
a quadcopter. A problem for existing fiducials is that low-cost
quadcopters often exhibit very quick and erratic physical movements
that result in motion blur in the captured images from the quadcopter's onboard
camera. This motion blur has an adverse effect on the recognition of fiducial
markers. This can be seen in Figure \ref{fig:teaser}--(b) where the
ARTag fiducial cannot be recognized due to motion blur. This is not
too surprising as most existing fiducials are not designed to handle
motion blur.

@@

{\textbf{Tracking:}}~~Fiducial detection between successive video
frames can be considered as a tracking problem where the tracked
object is the fiducial.  There is a very large body of research
dedicated to tracking and interested readers are referred
to~\cite{Yilmaz:2006} for a good survey.

@@

Here we examine the performance of two popular fiducials under
motion blur.  Specifically we examine ARTags ~\cite{Fiala05} and
PiTags~\cite{Pitag13} given their differences in geometric design and
the availability of an API to develop applications to recognize the tags.
\footnote{RUNE-tag~\cite{runetag11} currently does not
provide access to an implementation to generate or recognize its tags. 
The circular data matrix~\cite{NaimarkF02} is available as a
commercial product, however it requires proprietary hardware.}

@@

%\noindent Details of each of the steps are as follows:\\
\textbf{Gabor filter:}~~A 2D Gabor filter is a Gaussian kernel function
modulated by a sinusoidal plane wave~\cite{Kruizinga:2002}. It is used to find
high gradient patches. In our case, it is used to detect portions
of the circular fiducials that were not affected by the blur.
We applied the Gabor filter in eight
different orientations ($\theta = 0, 45, 90, \ldots, 225, 270, 315$).  The
following parameters were used for creating each Gabor kernel: $\lambda$ (wavelength) $= 8$, $\gamma$
(aspect ratio) $= 0.5$, $\sigma$ (spread) $= 0.56\lambda$, $\psi$
(phase angle) $= 0$ (for real part), $\pi/2$ (for imaginary part).
Then $\ell^2$-norm of outputs along all orientations is calculated and finally
$\ell^2$-normalized image is thresholded with threshold set to $0.4$ (on scale of zero to
one).  For further details to the Gabor filter, see~\cite{Kruizinga:2002}.
%% MBS - is the Kruizinga filter reasonable to cite?

\textbf{Clustering:}~~The binarized Gabor filter responses are
treated as a set of connected components in the image.   Clustering
is used to find components that are located in a close spatial region.  We do
this via hierarchical clustering~\cite{ALGLIB} using unweighted
average linkage with a distance threshold set to 150.

\textbf{PCA:}~~For each clustered set of components, we apply
PCA to determine the dominant direction.  This is done by examining
the orientation of the first principal component.  We then extract an
intensity profile patch in the input image along this direction as it
extends through the bounding box of the cluster. The signature of this
profile will be used to identify the code.

@@

\textbf{Classification:}~~As mentioned above, a small image patch
containing the intensity profile of the fiducial is used to identify
which of the many fiducials might be present and is represented by a
code.  We found that a training-based method using the k nearest
neighbor (k-NN) technique gave better results than trying to find
binary code directly by determining the presence (or absence) of ring
at particular positions in the image patch. This required
training-data which was easy to generate. A synthetic fiducial
is blurred along 36 orientations (0, 10, 20, \ldots , 350) with blur
scale set to 40, and the intensity profile along the first principal
component from every output is taken as training data for that
fiducial. Figure~\ref{fig:training_data} shows the process of
creating training data for fiducial with binary code ``01'' embedded
in it.  For a query image patch, we normalize the intensity range, and
then compare this information against the training data. The class
label from the closest top $K=5$ images in the training data is used
to label the patch.

@@

Our system has also been tested on images containing multiple fiducial
patterns in the same frame. Our algorithm successfully detected
all fiducials as well as correctly classified them as shown in
Figure \ref{fig:output_all}.

@@

Qualitatively we see in Figure~\ref{fig:blur_maximum} that despite of
even more blur than earlier described (50 units or more) detection is
still feasible and practical.  Quantitatively, the comparison of
recognition rate is shown in Figure~\ref{fig:recognition_rate}.
Even under more severe blur, of 65 pixels, the algorithm is
successful.  Our recognition is 100\% for all codes except the ``00''
which is undetectable after 50 pixels blur (which is still
significantly better than ARTag).  Reasons for the lower performance
when``00'' tags are present is discussed in
Section~\ref{sec:discussion}.

@@

In another setup we arranged our fiducials on four sides of a box and
revolved the quadcopter around the box. Later, we repeated the process
by replacing our fiducial by ARTags. Figure~\ref{fig:setup} shows some
frames from this sequence as well as the performance of both
fiducials. The overall detection rate of our tags is 90\% while that
of ARTags is only around 60\%.

@@

We also compare our approach with tracking designed for blurred input
scenes.  We have used four image sequences (consisting of around 1000
frames each). Each sequence contains different fiducials so that we
are able to contrast the performance of BLUT~\cite{Wu:2011} with the
proposed method.  From the images in the top rows of
Figure~\ref{fig:BLUT_compare_00} and Figure~\ref{fig:BLUT_compare_01},
we can see that BLUT is able to track the fiducial when the position
of fiducial does not change too much in successive frames. Also, it
can be seen that once BLUT loses track of the fiducial, it is not
able to recover. Since our approach detects the code in each frame,
large changes in the pattern's position is not an issue.  Some of our
detection results are also shown in Figure~\ref{fig:BLUT_compare_00}
and Figure~\ref{fig:BLUT_compare_01}.

@@

We found that sometimes our detection algorithm fails to recognize the
``00'' fiducial, when there is severe blur.  This is because when there
is too much blur, the innermost ring's response in the Gabor output is
too low and not detected properly.  This problem can be resolved by
increasing the radius of innermost ring to reduce the effect of blur
on the innermost ring.
%\noindent\textbf{Pose Estimation}~~

@@

In terms of numbers of fiducials, we are able to generate less markers
than ARTag. Many applications in robotics (e.g., quadcopter navigation)
may not require simultaneous use of a large number of fiducials.  Most
of the time, it is sufficient to have 4-6 different
fiducials. Nevertheless, we may be able to generate a larger number of
fiducials by using color backgrounds. This is an area of interest for
further research.
